---
title: "Final Tutorial"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=TRUE, error=TRUE)
```

  The use of analytical statisical data can prove to be one of the most relevant parts of research. Using an option such as GGplot2 can enhance data to the extent an individual without a strong scientific background can easily comprehend it. Data visualtization within GGplot2 can consist of such elements as: 
  Violin Plots - geom_violin()
```{r, echo=TRUE} 
ggplot(mpg, aes(drv, hwy)) + geom_violin()
```
  Box Plots - geom_boxplot()
```{r, echo=TRUE} 
ggplot(diamonds, aes(carat, depth)) +
  geom_boxplot(aes(group = cut_width(carat, 0.1))) +
  xlim(NA, 2.05)
```
  Scatter Plots - geom_point()
```{r, echo=TRUE} 
  ggplot(mpg, aes(displ, hwy, colour = factor(cyl))) +
  geom_point()
```
  Bar Graphs - geom_bar()
```{r, echo=TRUE} 
dplot <- ggplot(diamonds, aes(color, fill = cut)) +
  xlab(NULL) + ylab(NULL) + theme(legend.position = "none")
dplot + geom_bar(position = "dodge")
```
  Surface Plots - geom_contour()
```{r, echo=TRUE} 
gplot(faithfuld, aes(eruptions, waiting)) +
 geom_contour(aes(z = density, colour = ..level..))
```

  GGplot2 can simply be implemented into Rstudio by library(ggplot2). The rest of the data visualization that occurs is included in the package. GG can be used tolayer the information by displaying data, creating a statistical summary, and showing metadata. 
  
  Displaying data can help show a detection of a pattern, the summary can assist in model predictions from the context of the data revealing subtlties, and metadata helps give meaning to the raw data itself in either the background or foreground. Such code as: geom_area(), geombar(), geom_line(), and geom_point() are basic variaitons of plots/graphs used to repesent data. 
  
  Creating custom ranges for the data represented in the plots is adjusted by settign the x and y values. Code shown below sets the limits in ranges, specific parameters can be controlled by the code shown below. 
 X/Y Ranges 
```{r, echo=TRUE} 
  df <- data.frame(x = 1:3, y = 1:3)
  base <- ggplot(df, aes(x, y)) + geom_point()
  print(base)
```  
 X/Y Parameters
```{r, echo=TRUE}  
  base + scale_x_continuous(limits = c(1.5, 2.5))
```  
  Aside from setting ranges on a plot you can also zoom into a plot to show a more specific data point relevant to the research. coord_cartesian adjusts the visualized X/Y limits of the data. For base mpg on the highway based on engine displacement can be respresented for all engine sizes or for the sake of independet research one may be interested in a smaller range of visualized data without altering the entire plot, simply by resettign the parameters of x. Each time "+" is utilized we are adding a layer or element to the plot. This can be shown with the code:
  
  Entire Plot:
```{r, echo=TRUE}   
  base <- ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_smooth()
```
  Zoomed In Plot:
```{r, echo=TRUE} 
  base + scale_x_continuous(limits = c(5, 7))
```
  Facet wrapping gives us the ability to present mutiple different variables or factors within an experiment. Laying out multiple plots into subsets visualizing data in individual panels. To represent the previos mpg data seperated by vehicle type and engine displacement we can use code such as:

```{r, echo=TRUE}   
  mpg2 <- subset(mpg, cyl != 5 & drv %in% c("4", "f") & class != "2seater")
  base <- ggplot(mpg2, aes(displ, hwy)) +
  geom_blank() +
  xlab(NULL) +
  ylab(NULL)
  base + facet_wrap(~class, ncol = 3)
```

  Knowing the fundamental grammar that goes into ggplot2 can expand the working knowledge and confidence of the user. Terms such as:
  