---
title: "Intrinsics"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 5
    #code_folding: hide
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, messages=TRUE, error=TRUE)
```

## List of Functions



###help(topic) - documentation on topic
```{r}

help(runif)

```

###?topic - id
```{r}




```

is() - show objects in the search path; specify pat="pat" to search on a pattern
```{r, eval=F}
a <- "troy"
is(a)? #not sure if this is right


```

dir() - show files in the current directory

```{r}


getwd() #what working directory i am using

dir() 
```

###list.files() - shows files in the directory

```{r}

list.files()

```
## INPUT AND OUTPUT

###save() - saves the specified objects (...) in the XDR platform independent binary format

```{r, eval=F}

save() #i picked files form list but it is saying they arent found 

```


###load() - load the datasets written with save

```{r, eval=F}

load(index.rmd) #says not found

```

###data() - loads specified data sets
```{r}


```

###library() - load add on packages from library
```{r}

library(bitops) #loading according, box became checked in system library

```

###read.table() - reads a file in table format and creates a data
frame from it; the default separator sep="" is any whitespace; use header=TRUE to read the first line as a header of column names; use as.is=TRUE to prevent character vectors from being converted to factors; use comment.char="" to prevent "#" from being interpreted as a comment; use skip=n to skip n lines before reading data; see the help for options on row naming, NA treatment, and others
```{r}

#???

```

###read.csv() - ("filename",header=TRUE) id. but with defaults set for reading comma-delimited files
```{r, eval=F}

myfile <- "troysdata.csv" #the data file i want to read
thedata <- read.csv(myfile, row.names=5)#only read row 5

print(thedata)

```

###scan() - function read data from screen or file. Large data can be scanned in by just copy and paste, for example paste from EXCEL. Then use "ctrl+v" to paste the data, the data type will be automatically determined.
```{r}
dir()
scan() #i dont see how this isnt found its clearly in the directory

```

###print() - prints its arguments; generic, meaning it can have different methods for different objects
```{r}

a <- "hello"
print(a)

```

###cat() - (..., file="", sep=" ") prints the arguments after coercing to character; sep is the character separator between arguments
```{r}
b <- c(1,2,3)
cat(b) #simply shows me whats in variabke (b)

```

###write.table() - prints x after converting to a data frame; if quote is TRUE,character or factor columns are surrounded by quotes ("); sep is the field separator; eol is the end-of-line separator; na is the string for missing values; use col.names=NA to add a blank column header to get the column headers aligned correctly for spreadsheet input
```{r, eval=F}
a <- data.frame(a=c(1,2,3),b=c(3,4,5)) #created a table

write.table(a,file="mytable.txt") #gave the file a name

#example list was saved in the notes document but i recently added it to the envirnment as a saved variable

```

##DATA CREATION 

###c() - combine arguements by forming a vector, if recursive=TRUE is used it will descend through lists combining all elements in one vector
```{r}

first_names <- c("asdasas", "fwefef", "sfweg")

write.table(first_names)

```

###from:to (where from and to are replaced with numbers, e.g. 1:10)
```{r}

a <- (1:5)
a
```

###seq() - generates a sequence by= specifies increment, length= specifies how long
```{r, eval=F}

seq(2:10, by = 2) # i esentially want to show only even numbers the by= isnt doing what i think it should 

seq(2:10, by = (1))
```


###rep() - used to replicate. replicate x times; use each= to repeat “each” element of x each times; rep(c(1,2,3),2) is 1 2 3 1 2 3; rep(c(1,2,3),each=2) is 1 1 2 2 3 3
```{r}

rep(c(1,2,3),each=2)

```


###data.frame - create a data frame of the named or unnamed arguments; data.frame(v=1:4,ch=c("a","B","c","d"),n=10);shorter vectors are recycled to the length of the longest
```{r}



```

###list() - create a list of the named or unnamed arguments
```{r}

a <- list(a=1,2,3,b="hello",d=2,3,4)
b <- list(a=2, b="hi", d=3)
print(a)
print(b)

```

###matrix() - build a matrix
```{r}

h <- matrix(1:20, nrow=5, ncol=5)

View(h)


```

###factor() - encodes a vector x as a factor
```{r}

sodium <- factor(c("low", "med", "high"))
levels(sodium)

sodium <- factor(sodium, levels=c("low", "med", "high"), ordered=TRUE)
levels(sodium)

min(sodium)
```

###rbind() - combine data/info by rows for matrices, data frames, etc. must have equivalent amount of rows, if not equate them by adding NA to either until same qty is reached
```{r}



```

###cbind() - combine data by columns same rules apply as rbind

```{r}



```

##Slicing and Extracting Data - indexing Vectors
### x[n] - what/remove is in this position in the matrix
```{r}

a = matrix( 
  c(2, 4, 3, 1, 5, 7), # the data elements 
   nrow=2,              # number of rows 
   ncol=3,              # number of columns 
   byrow = TRUE)        # fill matrix by rows 
 
a                      # print the matrix 

a[2 , 1]   # first thing in second row

```

###x[-n] - everything but what is in the [] position
```{r}

a = matrix( 
  c(2, 4, 3, 1, 5, 7), # the data elements 
   nrow=2,              # number of rows 
   ncol=3,              # number of columns 
   byrow = TRUE)        # fill matrix by rows 
 
a                      # print the matrix 

a[-2, ] # everything but the first thing in the second row

```

### x[1:n] what is in positions 1-n
```{r}

a = matrix( 
  c(2, 4, 3, 1, 5, 7), # the data elements 
   nrow=2,              # number of rows 
   ncol=3,              # number of columns 
   byrow = TRUE)        # fill matrix by rows 
 
a                      # print the matrix 

a[1:2, c(1,1,1,1,1)] # everything but the first thing in the second row

```

### x[-(1:n)] - elements from n+1 to the end
```{r}

a = matrix( 
  c(2, 4, 3, 1, 5, 7),  # the data elements 
   nrow=2,              # number of rows 
   ncol=3,              # number of columns 
   byrow = TRUE)        # fill matrix by rows 
 
a [-(1:3)]              # print the matrix 


```

### x[c(1,4,2)] - specific elements
```{r}
a = matrix( 
  c(2, 4, 3, 1, 5, 7),  # the data elements 
   nrow=2,              # number of rows 
   ncol=3,              # number of columns 
   byrow = TRUE)        # fill matrix by rows 

a[c(1,3,5)] #what is in postion 1,3,5


```

### x["name"]
```{r, eval=FALSE}

k <- matrix(1:9, nrow = 3 dimnames = list( 
c("row1", "row2"),         # row names 
c("col1", "col2", "col3")) # column names 
 
k      #print A 

 
 # element at 2nd row, 3rd column 


```

###x[x>3] all elements greater than 3 - 
```{r}

x <- c(1,3,4,5,6,7)
x[x > 3] 
```

###x[x > 3 & x < 5] all elements between 3 and 5
```{r}

x <- c(1,3,4,5,6,7)
x[x > 3 & x < 5] 


```

###x[x %in% c(“a”,“and”,“the”)] all elements in given set
```{r}




```

##Indexing lists

###x[n] list with elements n
```{r}

n <- c(2, 3, 5) 
s <- c("aa", "bb", "cc", "dd", "ee")
b <- c(TRUE, FALSE, TRUE, FALSE, FALSE) 
x <- list(n, s, b, 3)   # x contains copies of n, s, b

x[4]

```

###x[[n]] nt h element of the list - 
```{r}

n <- c(2, 3, 5) 
s <- c("aa", "bb", "cc", "dd", "ee")
b <- c(TRUE, FALSE, TRUE, FALSE, FALSE) 
x <- list(n, s, b, 3)   # x contains copies of n, s, b

x[[1]][2]

```

###x[[“name”]] element of the list named “name”
```{r}

b[[TRUE]]

```

###x$name id.
```{r}
bob <- c(2, 3, 5)
john <- c("aa", "bb")
v <- list(bob <- c(2, 3, 5), john <- c("aa", "bb")) 
 
v$bob #i expect this to return 2,3,5
v[bob]

```

##Indexing matrices

###x[y,b] element at row y, column b
```{r}

x <- matrix(1:9, nrow = 3, dimnames = list(c("X","Y","Z"), c("A","B","C")))
x

x["Y","B"]
```

###x["Z",] row Z
```{r}

x["Z",]

```

###x[,c] column 
```{r}

x[,"C"]

```

###x[,c("A","C")] columns A and C
```{r}

x[,c("A","C")]

```

###x[“name”,] row named “name”
```{r}

x["Y",]

```

##Indexing data frames (matrix indexing plus the following)

###x[[“name”]] column named “name”
```{r, eval=F}

x <- matrix(1:9, nrow = 3, dimnames = list(c("X","Y","Z"), c("A","B","C")))
x[["B"]]

```

###x$named
```{r}



```

##Variable conversion
###as.data.frame(x) - 
```{r, eval=F}

as.data.frame(x, row.names=NULL, optional=FALSE, xy=FALSE, 
              na.rm=FALSE, long=FALSE, ...)#not sure exactly what i am attempting to do here

```

###as.numeric(x) - 
```{r}

x <- c(1,2,4,5,53,2)
as.numeric(x)# i believe this is should be coding responses as numeric, which i dont quite undetstand in this context.for instance a likert scale 

```

###as.logical(x)
```{r}


```

###as.character(x)
```{r}


```

##Variable information

###is.na(x)
```{r}


```

###is.null(x)
```{r}


```

###is.data.frame(x)
```{r}



```

###is.numeric(x)
```{r}


```

###is.character(x)
```{r}


```

###length(x)
```{r}

x <- c(1,2,5,6,3,6,64,4,3,7,8,9)
length(x)

```

###dim(x)
```{r}
x <- c(1,2,5,6,3,6,64,4,3,7,8,9)
dim(x)#? not sure here
```

###dimnames(x)
```{r}


```

###nrow(x)ncol(x) - amount of col/rows, also to use name 
```{r}
matrix(a, nrow = 2, ncol = 3)

```

###class()
```{r}


```

###attributes()
```{r}


```

##Data selection and manipulation

###which.max() - returns the largest number
```{r}

x <- c(1,4,57,54754,6576,3456346,4)

which.max(x) #not sure why its not diagnosing thr enetire number and just the last digit, 

```

###which.min() - smallest element of 
```{r}

x <- c(1211,124,57,54754,6576,3456346,4)
which.min(x) #why is this not 4?

```

###which() - what postion something is in
```{r}
y <- c(3,5,7,8,4,3,3,2)
which(y == 4)

```

###sort() - puts things in order
```{r}

m <- c(5,2,6,7,8,3,3,3,3)
sort(m)

```

###unique()
```{r}


```

###table() - can be used for frequency tables
```{r}
d <- c(2,4,5,8,5,3,6,5,8,6,5,3,3,4,6,6,8,8,5,4,3,2,3)
table(d)#, or categorization, assuming i can use strsplit but im not positive


```
###sample - simply put to sample numbers, but you can modify to sample greater or less then and incorporate sort to clean it up
```{r}

d <- 1:20
sort(sample(d[d > 7]))#sorted all numbers greater than 7 from variable d

```

##MATH

###max() - largest number in variable
```{r}

max(d)

```

###min() - smallest number in variable
```{r}
min(d)
```

###range() - from smallest to largest number
```{r}
range(d)

```

###sum - sum of all integers of a given variable
```{r}

sum(d)
```
###mean 
```{r}
mean(d)
```

###median
```{r}
median(d)
```

###var - The average of the squared differences from the Mean
```{r}

var(d)

```

###sd - Standard Deviation is just the square root of Variance
```{r}
sd(d)

```

###cor - 
```{r, eval=F}
n <- c(2,2,4,4,6,6,6,3,3,3,3)
cor(d,n)#just not sure here abount what to do

```

###round() - specify whole or decimal numbers in a sequence or within a variable
```{r}
round(.5 + -2:4)
( x1 <- seq(-2, 4, by = .5) )#sequence of numbers by the .5 in each sequence 
```

###abs() - 
```{r}

abs(d)
x <- matrix(c(-3,5,-7,1,-9,4),nrow=3,ncol=2,byrow=TRUE)
abs(x[1,])#absolute value of row one


```

##Matrices

###t()
```{r}

h <- matrix(1:20, nrow=5, ncol=4)
t(h)

```

###diag() - not sure what this is
```{r}

# NOT RUN {
dim(diag(3))
diag(10, 3, 4) # guess what?
all(diag(1:3) == {m <- matrix(0,3,3); diag(m) <- 1:3; m})

```

###rowsums() / colsums() - sum of rows and columns. rowmeans() / Colmeans()
```{r}
h <- matrix(1:20, nrow=5, ncol=4)
rowSums(h)
colSums(h)
rowMeans(h)
colMeans(h)
```

##Advanced data processing
###apply
```{r, eval=F}
apply

```

###aggregate()
```{r}

#not sure why or what this is going to do for me 
```

##Strings

###paste()
```{r}

paste0(1:12) #need an elementary explanation
paste(1:12)      
as.character(1:12)

```

###strsplit - strsplit helps find letters in string
```{r}

q <- "dogs cats" #strsplit helps find letters in string
strsplit(q,split="") #to split a character vector
t <-strsplit(q,split ="")



t[[1]]
t[[1]][2]#use to access second item in string
t[[1]][6]#use to access sixth item in string
table(t)



```

###tolower() - 
```{r}

test1 <- c(text1 = "England and France are members of NATO and UNESCO", 
text2 = "NASA sent a rocket into space.")
tolower(test1)

```

###toupper() - 
```{r}

test1 <- c(text1 = "England and France are members of NATO and UNESCO", 
text2 = "NASA sent a rocket into space.")
toupper(test1)

```

##Plotting

###hist()
```{r}

v <- c(3, 4, 7, 6, 5, 1, 4, 3, 4, 4, 5, 4, 0, 3, 5, 4, 4, 4)

# Give the chart file a name.
png(file = "histogram_lim_breaks.png")

# Create the histogram.
hist(v,xlab = "Weight",col = "green",border = "red", xlim = c(0,8), ylim = c(0,10),
   breaks = 5)

```

###plot()
```{r}

v <- c(3, 4, 7, 6, 5, 1, 4, 3, 4, 4, 5, 4, 0, 3, 5, 4, 4, 4)

# Give the chart file a name.
png(file = "histogram_lim_breaks.png")

# Create the histogram.
plot(v,xlab = "Weight",col = "green", xlim = c(0,8), ylim = c(0,10))

```

##Distributions

###rnorm - generates random deviates.
```{r}

rnorm(n, mean = 10, sd = 1.5)

```
###runif - random numbers
```{r}
n <- c(1,2,5,7,8,4,3,2,5,6,8,8,5)
runif(n, min = 0, max = 10)

```

##Programming

###show that you can define a function
```{r}

addtwo <- function(x) {
  b <- x + 2
  return(b)
}
addtwo(2) 

sumsq <- function(x,y) {
  x^2 + y^2
}
sumsq(2,3)#i think i need to insert a table to truly make this work 

```

###show that you can write a for loop
```{r}



```

###show that you can write a while loop
```{r}

x <- 4
while(x < 100) {
  x <- x + 8

print(x)
}
```

###show that you can write an if else statement
```{r}


j = c(1:30)
ifelse(j %% 3 == 0, "three", "not")

```

###Explain what return() does inside a function, show you can use it
```{r}
check <- function(x) {
if (x > 0) {
result <- "Positive"
}
else if (x < 0) {
result <- "Negative"
}
else {
result <- "Zero"
}
return(result)
}
check(1)
check(-3)
check(0)

```

###Explain what break() does, show you can use it
```{r}



```