---
title: "Intrinsics"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## List of Functions



###help(topic) - documentation on topic
```{r}

help(runif)

```

###?topic - id
```{r}




```

is() - show objects in the search path; specify pat="pat" to search on a pattern
```{r, eval=F}
a <- "troy"
is(a)? #not sure if this is right


```

dir() - show files in the current directory

```{r}


getwd() #what working directory i am using

dir() 
```

###list.files() - shows files in the directory

```{r}

list.files()

```
## INPUT AND OUTPUT

###save() - saves the specified objects (...) in the XDR platform independent binary format

```{r, eval=F}

save() #i picked files form list but it is saying they arent found 

```


###load() - load the datasets written with save

```{r, eval=F}

load(index.rmd) #says not found

```

###data() - loads specified data sets
```{r}


```

###library() - load add on packages from library
```{r}

library(bitops) #loading according, box became checked in system library

```

###read.table() - reads a file in table format and creates a data
frame from it; the default separator sep="" is any whitespace; use header=TRUE to read the first line as a header of column names; use as.is=TRUE to prevent character vectors from being converted to factors; use comment.char="" to prevent "#" from being interpreted as a comment; use skip=n to skip n lines before reading data; see the help for options on row naming, NA treatment, and others
```{r}

#???

```

###read.csv() - ("filename",header=TRUE) id. but with defaults set for reading comma-delimited files
```{r, eval=f}

myfile <- "troysdata.csv" #the data file i want to read
thedata <- read.csv(myfile, row.names=5)#only read row 5

print(thedata)

```

###scan() - function read data from screen or file. Large data can be scanned in by just copy and paste, for example paste from EXCEL. Then use "ctrl+v" to paste the data, the data type will be automatically determined.
```{r}
dir()
scan() #i dont see how this isnt found its clearly in the directory

```

###print() - prints its arguments; generic, meaning it can have different methods for different objects
```{r}

a <- "hello"
print(a)

```

###cat() - (..., file="", sep=" ") prints the arguments after coercing to character; sep is the character separator between arguments
```{r}

cat(b) #simply shows me whats in variabke (b)

```

###write.table() - prints x after converting to a data frame; if quote is TRUE,character or factor columns are surrounded by quotes ("); sep is the field separator; eol is the end-of-line separator; na is the string for missing values; use col.names=NA to add a blank column header to get the column headers aligned correctly for spreadsheet input
```{r}


write.table(example_list) #example list was saved in the notes document but i recently added it to the envirnment as a saved variable

```

##DATA CREATION 

###c() - combine arguements by forming a vector, if recursive=TRUE is used it will descend through lists combining all elements in one vector
```{r}

first_names <- c("asdasas", "fwefef", "sfweg")

write.table(first_names)

```

###from:to (where from and to are replaced with numbers, e.g. 1:10)
```{r}

a <- (1:5)
a
```

###seq() - generates a sequence by= specifies increment, length= specifies how long
```{r}

seq(2:10, by = 2) # i esentially want to show only even numbers the by= isnt doing what i think it should 

seq(2:10, by = (1))
```


###rep() - used to replicate. replicate x times; use each= to repeat “each” element of x each times; rep(c(1,2,3),2) is 1 2 3 1 2 3; rep(c(1,2,3),each=2) is 1 1 2 2 3 3
```{r}

rep(c(1,2,3),each=2)

```


###data.frame - create a data frame of the named or unnamed arguments; data.frame(v=1:4,ch=c("a","B","c","d"),n=10);shorter vectors are recycled to the length of the longest
```{r}



```

###list() - create a list of the named or unnamed arguments
```{r}

a <- list(a=1,2,3,b="hello",d=2,3,4)
b <- list(a=2, b="hi", d=3)
print(a)
print(b)

```

###matrix() - build a matrix
```{r}

h <- matrix(1:20, nrow=5, ncol=5)

View(h)


```

###factor() - encodes a vector x as a factor
```{r}

sodium <- factor(c("low", "med", "high"))
levels(sodium)

sodium <- factor(sodium, levels=c("low", "med", "high"), ordered=TRUE)
levels(sodium)

min(sodium)
```

###rbind() - combine data/info by rows for matrices, data frames, etc. must have equivalent amount of rows, if not equate them by adding NA to either until same qty is reached
```{r}



```

###cbind() - combine data by columns same rules apply as rbind

```{r}



```

##Slicing and Extracting Data - indexing Vectors
### x[n] - what/remove is in this position in the matrix
```{r}

a = matrix( 
  c(2, 4, 3, 1, 5, 7), # the data elements 
   nrow=2,              # number of rows 
   ncol=3,              # number of columns 
   byrow = TRUE)        # fill matrix by rows 
 
a                      # print the matrix 

a[2] # first thing in second row

```

###x[-n] - everything but what is in the [] position
```{r}

a = matrix( 
  c(2, 4, 3, 1, 5, 7), # the data elements 
   nrow=2,              # number of rows 
   ncol=3,              # number of columns 
   byrow = TRUE)        # fill matrix by rows 
 
a                      # print the matrix 

a[-2,2] # everything but the first thing in the second row

```

### x[1:n] what is in positions 1-n
```{r}

a = matrix( 
  c(2, 4, 3, 1, 5, 7), # the data elements 
   nrow=2,              # number of rows 
   ncol=3,              # number of columns 
   byrow = TRUE)        # fill matrix by rows 
 
a                      # print the matrix 

a[1:3] # everything but the first thing in the second row

```

### x[-(1:n)] - elements from n+1 to the end
```{r}

a = matrix( 
  c(2, 4, 3, 1, 5, 7),  # the data elements 
   nrow=2,              # number of rows 
   ncol=3,              # number of columns 
   byrow = TRUE)        # fill matrix by rows 
 
a [-(1:3)]              # print the matrix 


```

### x[c(1,4,2)] - specific elements
```{r}
a = matrix( 
  c(2, 4, 3, 1, 5, 7),  # the data elements 
   nrow=2,              # number of rows 
   ncol=3,              # number of columns 
   byrow = TRUE)        # fill matrix by rows 

a[c(1,3,5)] #what is in postion 1,3,5


```

### x["name"]
```{r, eval=F}

dimnames(A) = list( 
c("row1", "row2"),         # row names 
c("col1", "col2", "col3")) # column names 
 
a      #print A 

 
a ["row2", "col3"] # element at 2nd row, 3rd column 


```