---
title: "The Struggle"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```

## Problem 1

Do simple math with numbers, addition, subtraction, multiplication, division

```{r}
1+1
3-2
4*6
5/7


```

## Problem 2

Put numbers into variables, do simple math on the variables
```{r}

troy <- 1
abc <- 389
a <- 1
b <- 2
c <- 3
a+c

b-troy+c


```

## Problem 3

Write code that will place the numbers 1 to 100 separately into a variable using for loop. Then, again using the seq function.

```{r}

for (i in 1:1011:20) {
  print(i)
}
for(i in 11:20) {
  print (i)
}


```
## Problem 4

Find the sum of all the integer numbers from 1 to 100.

```{r}

sum(1:100)



```

## Problem 5
Write a function to find the sum of all integers between any two values.

```{r}



```
## Problem 6

List all of the odd numbers from 1 to 100.

```{r}

for (i in 1:99) {
if (!i %% 2){
    next
    }
  print(i)
}  


```

## Problem 7

List all of the prime numbers from 1 to 1000.
```{r}



```


## Problem 8

Generate 100 random numbers

```{r}

the_numbers <- runif (100, min=1, max=200)
the_numbers

```

## Problem 9

Generate 100 random numbers within a specific range

```{r}
the_numbers <- runif (100, min=1, max=100)
the_numbers 

```

## Problem 10

Write your own functions to give descriptive statistics for a vector variable storing multiple numbers. Write functions for the following without using R intrinsics: mean, mode, median, range, standard deviation

```{r}
some_numbers <- c(2,3,4,5,5,6,8) #c()=dataset
mean(some_numbers) #finding mean
median(some_numbers) #median
range(some_numbers) #range
sd(some_numbers) #standarddeviation
y <- table(some_numbers)#finding mode
y
names(y)[which(y==max(y))]

```

## Problem 11

Count the number of characters in a string variable

```{r}

q <- "dogs cats" #strsplit helps find letters in string
strsplit(q,split="")
t <-strsplit(q,split ="")

t[[1]]
t[[1]][2]#use to access second item in string
t[[1]][6]#use to access sixth item in string
```


## Problem 12

Count the number of words in a string variable
```{r}

a <- "counting words in a string"
strsplit(a,split=" ") #Error in strsplit(a, split = " ") : object 'a' not found but still allows to knit and send through gateway to website


```
