---
title: "Journa"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```

## Problem 1

Do simple math with numbers, addition, subtraction, multiplication, division

```{r}
1+1
3-2
4*6
5/7


```

## Problem 2

Put numbers into variables, do simple math on the variables
```{r}

troy <- 1
abc <- 389
a <- 1
b <- 2
c <- 3
a+c

b-troy+c


```

## Problem 3

Write code that will place the numbers 1 to 100 separately into a variable using for loop. Then, again using the seq function.

```{r} 

# i want it to do rows but i am getting stuck with one long column

for (i in 1:100) {
  print(i)
}

```
## Problem 4

Find the sum of all the integer numbers from 1 to 100.

```{r}

sum(1:100)



```

## Problem 5
Write a function to find the sum of all integers between any two values.

```{r}
a <- 5
b <- 10

sum(a:b)

```
## Problem 6

List all of the odd numbers from 1 to 100.

```{r}

for (i in 1:99) {
if (!i %% 2){
    next
    }
  print(i)
}  


```

## Problem 7

List all of the prime numbers from 1 to 1000.
```{r}



```


## Problem 8

Generate 100 random numbers

```{r}

the_numbers <- runif (100, min=1, max=200)
the_numbers

```

## Problem 9

Generate 100 random numbers within a specific range

```{r}
the_numbers <- runif (100, min=1, max=100)
the_numbers 

```

## Problem 10

Write your own functions to give descriptive statistics for a vector variable storing multiple numbers. Write functions for the following without using R intrinsics: mean, mode, median, range, standard deviation

```{r}
some_numbers <- c(2,3,4,5,5,6,8) #c()=dataset
mean(some_numbers) #finding mean
median(some_numbers) #median
range(some_numbers) #range
sd(some_numbers) #standarddeviation
y <- table(some_numbers)#finding mode
y
names(y)[which(y==max(y))]

```

## Problem 11

Count the number of characters in a string variable

```{r}

q <- "dogs cats" #strsplit helps find letters in string
strsplit(q,split="")
t <-strsplit(q,split ="")

t[[1]]
t[[1]][2]#use to access second item in string
t[[1]][6]#use to access sixth item in string
```


## Problem 12

Count the number of words in a string variable
```{r}

a <- "counting words in a string"
strsplit(a,split=" ") 


```

## Problem 13  ????

Count the number of sentences in a string variable

```{r}

a <- "coding is hard. but i like the challenge"
  strsplit(a,split="") #not sure how to count sentences


```


## Problem 14

14.	Count the number of times a specific character occurs in a string variable

```{r}


a <- c(2,5,3,4,5,7,5,4,3,3,5,6,8,9,0,7,4,3,2,3,5,8,6,4)
table(a) 


```

## Problem 15  ????

Do a logical test to see if one word is found within the text of another string variable.

```{r}

test_word <- "hi"  
test_sentence <- "hi my name is troy"

test_word%in%test_sentence #should be true...

c <- c(1,2,3,4,5,6)
d <- 2
e <- 7

d%in%c
e%in%c
```

##Problem 16

Put the current computer time in milliseconds into a variable

```{r}

print(as.numeric(Sys.time())*1000, digits=15)




```

## Problem 17

Measure how long a piece of code takes to run by measuring the time before the code is run, and after the code is run, and taking the difference to find the total time

```{r}



```
