---
title: "Statistics"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 5
    #code_folding: hide
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=TRUE, error=TRUE)
```


#Samples
###one direction t test
```{r}

x <- c(1,4,3,2,3,4,3,2,3,2,3,4)

#non-directional
t.test(x, mu = 2)


```
###paired sample t test
```{r}

x <- c(1,4,3,2,3,4,3,2,3,2,3,4)
y <- c(3,2,5,4,3,2,3,2,1,2,2,3)

#non-directional
t.test(x, y, paired=TRUE)


```

### independent samples t test
```{r}
x <- c(1,4,3,2,3,4,3,2,3,2,3,4)
y <- c(3,2,5,4,3,2,3,2,1,2,2,3)

#non-directional
my_ttest <- t.test(x, y, var.equal=TRUE)

my_ttest$p.value #using $ sign gets specific values

#

```

### fake paired t test
```{r}
 
a <- replicate(10000,t.test(rnorm(10,0,1),rnorm(10,0,1)))statistic

a

```

###printing a clean table 
```{r}

library(broom)

t_results <- t.test(x, y, var.equal=TRUE)
knitr::kable(tidy(t_results))


```



##ANOVA
### anova template for 2x2
```{r}
Subjects <- rep(1:10,each=4)
Factor1 <- rep(rep(c("A","B"), each = 2), 10)
Factor2 <- rep(rep(c("1","2"), 2), 10)
DV <- rnorm(40,0,1)

all_data <- data.frame(Subjects = as.factor(Subjects),
                       DV,
                       Factor1,
                       Factor2)



```

###between-subjects 1 factor template
```{r}

factor1 <- rep(c("A","B","C"),each = 5) #does each factor 5 times
DV <- rnorm(n=15,mean=0,sd=1)

all_data <- data.frame(DV,factor1)

#run the anova
aov_out <- aov(DV~Factor1, all_data) #dv "~" name of factor

#all the details of the anova in a summary table
summary(aov_out) 



```

###between-subjects 2 factor template
```{r}
# run anova
aov_out <- aov(DV~Factor1*Factor2, all_data)

# summary
summary(aov_out)



```

###repeated-measures 1 factor template - when the subject does each factor 
```{r}

# run anova - error is the name of the variable that we must divide by the factor
aov_out <- aov(DV~Factor1 + Error(Subjects/Factor1), all_data)

# summary
summary(aov_out)

```


###repeated-measures 2 factor
```{r}

# run anova
aov_out <- aov(DV~Factor1*Factor2 + Error(Subjects/(Factor1*Factor2)), all_data)

# summary
summary(aov_out)


# kable xtable printing
library(xtable)#library "()" means what package
knitr::kable(xtable(summary(aov_out)))#allows me to load a function within a package without loading a package

sum_out[[1]]$'F value'
```

##Linear Regression
```{r}
#lm is linear regression
lm(DV~Factor1, all_data)



```

#Troy's Stat Homework

##Null distribution of mean differences

```{r}



k <- replicate(10000,mean(rnorm(10,0,1)))-mean(rnorm(10,0,1))

data.frame(k)

ggplot(aes(x = k)) + geom_histogram() + theme_bw()




```


##t-distribution
```{r}




k <- replicate(10000,t.test(rnorm(10,0,1),mu=0)$statistic)

length(k[k < .5])/10000
```

##Correlation
```{r}

e <- rnorm(10,0,1)
f <- rnorm(10,0,1)


cor(e,f)

```

##F-values
```{r}

A <- c(1,2,3,4)
B <- c(3,4,5,6)
C <- c(5,6,7,8)


```

##F simulation
```{r}

L <- replicate(10000(rnorm(10,0,1)))


```

HW

#Correlataion

```{r}
#create a sampling distribution for correlations we can get in a specific situation

cor(rnorm(100,0,1),rnorm(100,0,1))#use cor function to correlate two sets

replicate(10000,cor(rnorm(100,0,1),rnorm(100,0,1)))

hist(replicate(1e+05,cor(rnorm(100,0,1),rnorm(100,0,1))))

sim_rs <- replicate(10000,cor(rnorm(100,0,1),rnorm(100,0,1)))

sort(abs(sim_rs))[95000]
```

#F simulation

```{r}
A <- c(1,2,3,4)
B <- c(3,4,5,6)
C <- c(5,6,7,8)

conds<- rep(c("A","B","C"), each=4)
DV <- c(A,B,C)
df <- data.frame(conds, DV)
summary (aov(DV~conds,df))
```

```{r}
run_anova <- function(){

A <- rnorm(4,0,1)
B <- rnorm(4,0,1)
C <- rnorm(4,0,1)

conds<- rep(c("A","B","C"), each=4)?
DV <- c(A,B,C)
df <- data.frame(conds, DV)
summary (aov(DV~conds,df))

sum_out <- summary(aov(DV~conds,df))
sum_out[[1]]$`F value`[1]

return(sum_out[[1]]$`F value`[1])
}


save_fs <- replicate(10000,run_anova())


hist(save_fs)

sort(save_fs)[9500]

qf(.95,2,9)
```