---
title: "Troy's R Basics"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, messages=TRUE, error=TRUE)
```

## Problem 1

Do simple math with numbers, addition, subtraction, multiplication, division

```{r}
1+1
3-2
4*6
5/7


```

## Problem 2

Put numbers into variables, do simple math on the variables
```{r}

troy <- 1
abc <- 389
a <- 1
b <- 2
c <- 3
a+c

b-troy+c


```

## Problem 3

Write code that will place the numbers 1 to 100 separately into a variable using for loop. Then, again using the seq function.

```{r} 

a <- length(100)
for (i in 1:100) { #: makes a sequence between numbers
  a[i] <-i
}
print (a)
```
## Problem 4

Find the sum of all the integer numbers from 1 to 100.

```{r}

sum(1:100)



```

## Problem 5
Write a function to find the sum of all integers between any two values.

```{r}
a <- 5
b <- 10

sum(a:b)

```
## Problem 6

List all of the odd numbers from 1 to 100.

```{r}
a <- length (1:99)
for (i in 1:99) {
if (!i %% 2){  
    next
    }
  print(i) #id like to print in rows
}  #!=not equal to i, %%=divided by

a <- 1:100
a[ a%%2 != 0]

```

## Problem 7
List all of the prime numbers from 1 to 1000.

```{r}

a <- length (1:1000)
for (i in 1:1000) {
if (!i %% 1)  
    next
     #id like to print in rows
}  #!=not equal to i, %%=divided by
print(i)
```

## Problem 8

Generate 100 random numbers

```{r}
a <- (1:200)

round(a, digits = 0) # not sure how to limit to 100 numbers

sample(1:200,200)
1:200
```


## Problem 9 

Generate 100 random numbers within a specific range

```{r}

round(runif(50, min = 100, max = 500))


```

## Problem 10

Write your own functions to give descriptive statistics for a vector variable storing multiple numbers. Write functions for the following without using R intrinsics: mean, mode, median, range, standard deviation
```{r}

some_numbers <- c(3, 4, 7, 6, 5, 1, 4, 3, 4, 4, 5, 4, 0, 3, 5, 4, 4, 4) #c()=dataset
mean(some_numbers) #finding mean
median(some_numbers) #median
range(some_numbers) #range
sd(some_numbers) #standarddeviation
y <- table(some_numbers)#finding mode
y
names(y)[which(y==max(y))]
```


## Problem 11

Count the number of characters in a string variable

```{r}

q <- "dogs cats" #strsplit helps find letters in string
strsplit(q,split="") #to split a character vector
t <-strsplit(q,split ="")



t[[1]]
t[[1]][2]#use to access second item in string
t[[1]][6]#use to access sixth item in string
table(t)
```


## Problem 12

Count the number of words in a string variable
```{r}

a <- "counting words in a string"
strsplit(a,split=" ") 
t <-strsplit(a,split ="") #not sure how to count the words instead of individual characters

table(t)

```

## Problem 13  ????

Count the number of sentences in a string variable

```{r}

a <- "coding is hard. but i like the challenge. cats. dogs"
b <- strsplit(a,split="[.]") #[.] sperates the sentences

length(b)
length(b[[1]]) #the amount of sentences in the vector 

length(a[[1]])

```


## Problem 14

14.	Count the number of times a specific character occurs in a string variable

```{r}


a <- c(2,5,3,4,5,7,5,4,3,3,5,6,8,9,0,7,4,3,2,3,5,8,6,4)
table(a) 


```

## Problem 15  ????

Do a logical test to see if one word is found within the text of another string variable.

```{r}

test_word <- "hi"  #"" is one thing
test_sentence <- "hi my name is troy"
unlist(strsplit(test_sentence, split = " ")) #flattens into a vector
f <- test_word #put string into variable
g <- test_sentence
a <- strsplit(test_sentence, split = " ")
b <- unlist(strsplit(test_sentence, split = " ")) #what does this mean
b[1]
a[[1]][1]#the first thing in the fist thing

test_word%in%f #because test word became variable instead of string we can now see if it is in f

f%in%a

f <- c(1,2,3,4,5,6)
d <- 2
e <- 7

d%in%f
e%in%f
```

##Problem 16

Put the current computer time in milliseconds into a variable

```{r, eval=FALSE}



print(as.numeric(Sys.time())*1000, digits=15)




```

## Problem 17

Measure how long a piece of code takes to run by measuring the time before the code is run, and after the code is run, and taking the difference to find the total time

```{r}
print(as.numeric(Sys.time())*10, digits=15)


```

## Problem 18

Read a .txt file or .csv file into a variable
```{r}




```


## Problem 19

Output the contents of a variable to a .txt file -write.csv()-
```{r}




```

## Problem 20

Create a variable that stores a 20x20 matrix of random numbers

```{r}

a <- matrix(1:500, ncol=20,nrow=20) 

  View(a)


```

## Problem 21

Output any matrix to a txt file using commas or tabs to separate column values, and new lines to separate row values

```{r}



```

#HARDER PROBLEMS

## Problem 1

The FizzBuzz Problem. List the numbers from 1 to 100 with the following constraints. If the number can be divided by three evenly, then print Fizz instead of the number. If the number can be divided by five evenly, then print Buzz instead of the number. Finally, if the number can be divided by three and five evenly, then print FizzBuzz instead of the number. The answer could look something like this:

```{r}
for (i in 1:100) {
 if (i %% 3 == 0 & i %% 5 == 0) {print("FizzBuzz")} #when used with curlys it will print keep in mind
 else if (i %% 3 == 0) {print("Fizz")}
 else if (i %% 5 == 0) {print("Buzz")}
 else print(i)
}



```

## Problem 2

Take text as input, and be able to produce a table that shows the counts for each character in the text. This problem is related to the earlier easy problem asking you to count the number of times that a single letter appears in a text. The slightly harder problem is the more general version: count the frequencies of all unique characters in a text

```{r}


k<-"the follow is text is written to count the individual letters"
table(unlist(strsplit(k,split="")))


```

```{r}
numbers <- c(1,6,4,2,3,4,5,5)
letters <- c("e","a","b","e","a","b","a","b")

new_df <- data.frame(numbers,letters)
print(new_df)
```

## Problem 3
###Test the random number generator for a flat distribution. Generate a few million random numbers between 0 and 100. Count the number of 0s, 1s, 2s, 3s, etc. all the way up to 100. Look at the counts for each of the numbers and determine if they are relatively equal. For example, you could plot the counts in Excel to make a histogram. If all of the bars are close to being flat, then each number had an equal chance of being selected, and the random number generator is working without bias.
```{r}

a<-runif(1000000,0,10000000)
hist(a)

```

## Problem 4
###Generate a matrix for a multiplication table. For example, the labels for the columns could be the numbers 1 to 10, and the labels for the rows could be the numbers 1 to 10. The contents of each of the cells in the matrix should be correct answer for multiplying the column value by the row value.
```{r}
#i cant for the life of me figure this problem out.....:(
a = matrix( 
  c(1:10),  # the data elements 
   nrow=10,              # number of rows 
   ncol=10,              # number of columns 
   byrow = TRUE)        
for(i in 1:10){
  for(j in 1:10){
    print(i*j)
  }
}
a

```

##Problem 5
###Encrypt and Decrypt the Alphabet Turn any normal english text into an encrypted version of the text, and be able to turn any decrypted text back into normal english text. A simple encryption would be to scramble the alphabet such that each letter corresponds to a new randomly chosen (but unique) letter.
```{r}

paste("1_", letters, sep = "")



```

##Problem 6
###Your task here is to write an algorithm that can simulate playing the above depicted Snakes and Ladders board. You should assume that each roll of the dice produces a random number between 1 and 6. After you are able to simulate one played game, you will then write a loop to simulate 1000 games, and estimate the average number of dice rolls needed to successfully complete the game.
```{r}



```
