---
title: "GGplot2"
subtitle: Visuals With GGplot2
author: "Troy Funderburk"
date: "2019/13/05 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---


```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

class: inverse, middle, center

# Different Options in GGplot2

- Violin Plot
- Surface Plots
- Bar Graph 
- Scatter Plots
- Box Plots

---

class: center, inverse

# How to Design and Alter a Violin Plot

- geom_violin()

```{r, echo=FALSE}
ggplot(mpg, aes(drv, hwy)) + geom_violin()
```

---

class: center, inverse

# How to make a 2D Image Look 3D With Surface Plots

-
```{r, echo=FALSE}

ggplot(faithfuld, aes(eruptions, waiting)) +
 geom_contour(aes(z = density, colour = ..level..))


```
-
```{r, echo=FALSE}

ggplot(faithfuld, aes(eruptions, waiting)) +
  geom_raster(aes(fill = density))

```

---

class: center, inverse
# Position Adjustments on Bar Graphs

- position_stack() stacks overlapping bars on top of eachother
- position_fill() stack bars so that the top is always number 1
- position_dodge() place bars/boxes side by side

-
```{r, echo=FALSE}

dplot <- ggplot(diamonds, aes(color, fill = cut)) +
  xlab(NULL) + ylab(NULL) + theme(legend.position = "none")

dplot + geom_bar()


```

---

class: center, inverse
# Position Fill Example

```{r, echo=FALSE}

dplot <- ggplot(diamonds, aes(color, fill = cut)) +
  xlab(NULL) + ylab(NULL) + theme(legend.position = "none")

dplot + geom_bar(position = "fill")


```

---

class: center, inverse
# Position Dodge Example

```{r, echo=FALSE}

dplot <- ggplot(diamonds, aes(color, fill = cut)) +
  xlab(NULL) + ylab(NULL) + theme(legend.position = "none")

dplot + geom_bar(position = "dodge")


```

---
class: center, inverse

# Box Plot Examples

## The box and whisker plots shows five summary statistics along with the individual outliers. Althought showing less than a histogram, it takes up far less space. 

```{r, echo=FALSE}

ggplot(diamonds, aes(carat, depth)) +
  geom_boxplot(aes(group = cut_width(carat, 0.1))) +
  xlim(NA, 2.05)

```

---

class: center, inverse

# Scatter Plot Examples

## In this example we can see how ggplot auto picks colors based on the different factors. 

```{r, echo=FALSE}

ggplot(mpg, aes(displ, hwy, colour = factor(cyl))) +
  geom_point()

```
---


